Profit Optimisation in Heterogeneous 2-link Parallel Games

### 2024-11-28 thesis/article.tex
Thesis: Generalises upper bound theorem for n-links

@@

\begin{lemma}
	\label{lemma:as_upper_bound}
	Let $([2], \ell, \alpha)$ a toll game with $x_1(0), x_2(0) > 0$.
	For all $t \in \R_+^2 \setminus \{(a, a) | a \in \R_+\}$, let .
	It holds that
	\[\as(t) \le \max\{\sup A_1^-(0), \sup A_2^-(0)\}\]
	or equivalently if $\alpha$ continuous at $x_1(0)$ and $x_2(0)$
	\[\as(t) \le \max\{\alpha(x_1(0)), \alpha(x_2(0))\}\]
\end{lemma}

\begin{proof}
	 From Lemmas \ref{lemma:split_asymptotic} and \ref{lemma:split_monotonicity} we get that $\as$, for $t_1 - t_2$ in $(-\infty, 0)$ and $(0, +\infty)$, is non-decreasing and non-increasing respectively.
	 Therefore $\as$ is upper bounded by its sided limits as $t_1 - t_2 \rightarrow 0$, which means that for all $t \in \R_+^2 \setminus \{(a, a) | a \in \R_+\}$ we get
	 \[\as(t) \le \max\left\{\lim_{t_1 - t_2 \rightarrow 0^-}\as(t), \lim_{t_1 - t_2 \rightarrow 0^+}\as(t)\right\}\]
	 which in turn after applying Lemma \ref{lemma:split_asymptotic}$(ii)$ becomes
	 \[\as(t) \le \max\{\sup A_2^-(0), \sup A_1^-(0)\}\]
\end{proof}

\begin{theorem}
	\label{theorem:alpha_upper_irrelevant}
	Let $([2], \ell, \alpha^{(1)})$ and $([2], \ell, \alpha^{(2)})$ two toll games with same latency functions, where $x_1(0), x_2(0) > 0$ for their common flow $x(0)$ and distribution functions $\alpha_1$ and $\alpha_2$ such that $\alpha^{(1)}(p) = \alpha^{(2)}(p), \forall p: p < \max\{x_1(0), x_2(0)\}$.
	Then the two games are equivalent.
\end{theorem}

\begin{proof}
	For tolls $t \in \R_+^2$ let $x^{(1)}(t), x^{(2)}(t)$ the Nash equilibria for $t$ for each game respectively.
	It suffices to show that $x^{(1)}(t) = x^{(2)}(t), \forall t \in \R_+^2$.
	It's already stated that the equality holds for $t = 0$, i.e. for $t \in \{(a, a) | a \in \R_+\}$.
	For the remaining $t$, assume w.l.o.g. that $t_1 > t_2$, which now makes it sufficient to show that the flow in the (common) lower split is equal, i.e $x_1^{(1)}(t) \ne x_1^{(2)}(t)$.

	Assume that $x_1^{(1)}(t) \ne x_1^{(2)}(t)$ and again w.l.o.g. that $x_1^{(1)}(t) < x_1^{(2)}(t)$.
	Then there exists a player $p$ such that $x_1^{(1)}(t) < p \le x_1^{(2)}(t)$ who is on link $2$ on the first game and on link $1$ for the second one.
	Also note that since $p$ is in the lower split, it holds that $p < \max\{x_1(0), x_2(0)\}$ so by the hypothesis we get $\alpha^{(1)}(p) = \alpha^{(2)}(p)$.
	With the above we get
	\[x_1^{(1)}(t) < x_1^{(2)}(t) \Rightarrow x_1^{(1)}(t) + \alpha^{(1)}(p)t_1 < x_1^{(2)}(t) + \alpha^{(2)}(p)t_1 \Rightarrow c_1^{p, (1)}(t) < c_1^{p, (2)}(t)\]
	and respectively
	\[x_2^{(1)}(t) > x_2^{(2)}(t) \Rightarrow x_2^{(1)}(t) + \alpha^{(1)}(p)t_2 > x_2^{(2)}(t) + \alpha^{(2)}(p)t_2 \Rightarrow c_2^{p, (1)}(t) > c_2^{p, (2)}(t)\]
	The above inequations mean that, between the first and the second game, for player $p$, the cost of link $1$ increased while the cost of link $2$ decreased, concluding that $p$ cannot be on link $2$ in the first but then change to link $1$ in the second, which is a contradiction.

	Therefore $x_1^{(1)}(t) = x_1^{(2)}(t)$ for all $t \in \R_+^2 \setminus \{(a, a) | a \in \R_+\}$, and in turn also $x^{(1)}(t) = x^{(2)}(t)$ for all $t \in \R_+$.
	To conclude, the two games are equivalent.
\end{proof}
We can see how this theorem applies to Example \ref{example:simple_alpha}, where since $x(0) = (2/3, 1/3)$ then $\max\{x_1(0), x_2(0)\} = 2/3$.
Then even if we used a distribution function such as
\[
	\alpha(p) =
	\begin{cases}
		p + 1 & p \le 2/3 \\
		p + 10 & p > 2/3
	\end{cases}
\]
then $\as$ would still behave the same way as those $(2/3, 1]$ players will always be in the upper split regardless of their $\alpha$ value.

It will become more apparent later when calculating best responses as even in the pricing games those higher values will never be used.
We finish this section with a lemma about the behaviour of $\as$ when many players have money sensitivity close to 0.
Note that it doesn't suffice to assume $\alpha(0) = 0$ as the behavior of that single player has no effect on the game, so we additionally assume that $0$ is an accumulation point in $\alpha$.
\begin{lemma}
	If for any $i = 1, 2 \; x_i(0) > 0$ and $\alpha(0) = 0$ with $0$ an accumulation point in $\alpha$, then $x_i(t) > 0, \forall t \in \R_+^2$
\end{lemma}

\begin{proof}
	By the definition of an accumulation point, we get for $0$ in $\alpha$ that $\forall \delta > 0 \; \exists \epsilon > 0: p < \epsilon \Rightarrow \alpha(p) < \delta$.
	Therefore $\forall t \in \R_+^2$ if $\delta = \as(t)$ then $\exists \epsilon > 0: p < \epsilon \Rightarrow \alpha(p) < \as(t)$, which means that the lower latency link will always have flow some $\epsilon > 0$, and therefore $x_1(t), x_2(t) > 0$.
\end{proof}

>>>>

TODO: Tolls define ordering $t_1 \ge t_2 \ge \dots t_n$ and $\ell_1(x_1(t)) \le \ell_2(x_2(t)) \le \dots \le \ell_n(x_n(t))$.

\begin{lemma}
	\label{lemma:xn_xn0_lower_bound}
	Let $(N, \ell, \alpha)$ a heterogeneous parallel game.
	For tolls $t \in \R_+^N$ consider the links in decreasing toll ordering.
	It holds that $x_n(t) \ge x_n(0)$.
\end{lemma}

\begin{proof}
	From assumption we have $t_n = \min_{i \in N} t_i$, which makes tolls $t^\prime = t - t_n$ be in $\R_+^N$ with $t_n^\prime = 0$.
	Since the flow equilibrium $x(t)$ is dependent only on the toll differences, we get $x(t) = x(t^\prime)$.
	Also, comparing $t^\prime$ with $t = 0$, no additional tolls have been added to link $n$, making $x_n(t^\prime) \ge x_n(0)$.
	The proposition follows naturally from the previous two statements.
\end{proof}

@@

\begin{lemma}
	\label{lemma:as_upper_bound}
	Let $(N, \ell, \alpha)$ a heterogeneous parallel game.
	For all $i, j \in N, i \ne j$ it holds that
	\[\as^{(i, j)}(t) \le \max\left\{\limsup_{t_i - t_j \rightarrow 0^+} A_i(t), \limsup_{t_j - t_i \rightarrow 0^+} A_j(t)\right\}\]
	%or equivalently if $\alpha$ continuous at $x_1(0)$ and $x_2(0)$
	%\[\as(t) \le \max\{\alpha(x_1(0)), \alpha(x_2(0))\}\]
\end{lemma}

\begin{proof}
	 From Lemmas \ref{lemma:split_asymptotic} and \ref{lemma:split_monotonicity} (also look the discussion at the function characterisation above) we get that $\as$, for $t_i - t_j$ in $(-\infty, 0)$ and $(0, +\infty)$, is non-decreasing and non-increasing respectively.
	 Therefore $\as$ is upper bounded by its sided limits as $t_i - t_j \rightarrow 0$, which means that for all $t$ in the domain of $\as$ we get
	 \[\as^{(i j)}(t) \le \max\left\{\lim_{t_i - t_j \rightarrow 0^+} \as^{(i, j)}(t), \lim_{t_i - t_j \rightarrow 0^-} \as^{(i, j)}(t)\right\}\]
	 which in turn, since $t_i - t_j \rightarrow 0^-$ is equivalent to $t_j - t_i \rightarrow 0^+$ and after applying Lemma \ref{lemma:split_asymptotic}$(ii)$, becomes
	 \[\as^{(i j)}(t) \le \max\left\{\limsup_{t_i - t_j \rightarrow 0^+} A_i(t), \limsup_{t_j - t_i \rightarrow 0^+} A_j(t)\right\}\]
\end{proof}

\begin{theorem}
	\label{theorem:alpha_upper_irrelevant}
	Let $\Gm_1 = (N, \ell, \alpha^{(1)})$ and $\Gm_2 = (N, \ell, \alpha^{(2)})$ be two heterogeneous parallel games with the common number of links $n = |N|$  and latency functions $\ell$.
	For $x^{(1)}(t), x^{(2)}(t)$ the Nash equilibria for $t \in \R_+^N$ in $\Gm_1, \Gm_2$ respectively, denote as $x(0) = x^{(1)}(0) = x^{(2)}(0)$ the common flow on link of $\Gm_1$ and $\Gm_2$ for $t = 0$.
	If the distribution functions $\alpha^{(1)}$ and $\alpha^{(2)}$ are such that $\alpha^{(1)}(p) = \alpha^{(2)}(p), \forall p: p < 1 - \min_{i \in N}\{x_i(0)\}$, then $\Gm_1$ and $\Gm_2$ are equivalent.
\end{theorem}

\begin{proof}
	It suffices to show that $x^{(1)}(t) = x^{(2)}(t), \forall t \in \R_+^N$.
	It's already stated that the equality holds for $t = 0$.
	%assume w.l.o.g. that $t_1 > t_2$, which now makes it sufficient to show that the flow in the (common) lower split is equal, i.e $x_1^{(1)}(t) \ne x_1^{(2)}(t)$.
	For $t \ne 0$, assume for simplicity that all $t_i \ne t_j$ for all $i, j \in N$.
	Note that since we apply the same tolls $t$ to both game, the ordering they impose on the links is the same.
	We denote as $i = i^{(1)} = i^{(2)}$ the common corresponding link $i$ between the two games.

	Assume that the theorem does not hold, i.e. $x^{(1)}(t) \ne x^{(2)}(t)$.
	Then there exists some link $i \in N$ such that $x_i^{(1)}(t) > x_i^{(2)}(t)$, and therefore there exists a player $p$ who is on link $i$ in $\Gm_1$ but not in $\Gm_2$.
	We can assume that for the link $j$ this player went to in $\Gm_2$ it holds $x_j^{(1)}(t) \le x_j^{(2)}(t)$; if not, then select $j$ and start the same reasoning from the beginning.

	If we consider now the split between $i$ and $j$, player $p$ exist in the lower split for one game and on the upper one for the other, let w.l.o.g. those game be $\Gm_1$ and $\Gm_2$ respectively.
	So if we take $\Gm_1$ where $p$ in the lower split, then it holds that $p$ is not in $x_n^{(1)}$, since the link with the highest toll is always the upper split with regard to all others.
	Therefore $p < 1 - x_n^{(1)}$ and from Lemma \ref{lemma:xn_xn0_lower_bound} we then get $p < 1 - x_n(0) < 1 - \min_{i \in N}\{x_i(0)\}$, which from the hypothesis leads us to $\alpha^{(1)}(p) = \alpha^{(2)}(p)$.

	With the above we get
	\[x_i^{(1)}(t) > x_i^{(2)}(t) \Rightarrow x_i^{(1)}(t) + \alpha^{(1)}(p)t_i > x_i^{(2)}(t) + \alpha^{(2)}(p)t_i \Rightarrow c_i^{p, (1)}(t) > c_i^{p, (2)}(t)\]
	and respectively
	\[x_j^{(1)}(t) \le x_j^{(2)}(t) \Rightarrow x_j^{(1)}(t) + \alpha^{(1)}(p)t_j \le x_j^{(2)}(t) + \alpha^{(2)}(p)t_j \Rightarrow c_j^{p, (1)}(t) \le c_j^{p, (2)}(t)\]
	The above inequations mean that, between $\Gm_1$ and $\Gm_2$, for player $p$, the cost of link $i$ decreased while the cost of link $j$ didn't, concluding that $p$ cannot be on link $i$ in $\Gm_1$ but then change to link $j$ in $\Gm_2$, which is a contradiction.

	Therefore $x^{(1)}(t) = x^{(2)}(t)$ for all $t \in \R_+$ and from Definition \ref{definition:game_equivalence} the two games are equivalent.
\end{proof}
We can see how this theorem applies to Example \ref{example:simple_alpha}, where since $x(0) = (2/3, 1/3)$ then $\max\{x_1(0), x_2(0)\} = 2/3$.
Then even if we used a distribution function such as
\[
	\alpha(p) =
	\begin{cases}
		p + 1 & p \le 2/3 \\
		p + 10 & p > 2/3
	\end{cases}
\]
then $\as$ would still behave the same way as those $(2/3, 1]$ players will always be in the upper split regardless of their $\alpha$ value.

It will become more apparent later when calculating best responses as even in the pricing games those higher values will never be used.
We finish this section with a lemma about the behaviour of $\as$ when many players have money sensitivity close to 0.
Note that it doesn't suffice to assume $\alpha(0) = 0$ as the behavior of a single player has no effect on the game, so we additionally assume that $0$ is an accumulation point in $\alpha$.
\begin{lemma}
	Let $(N, \ell, \alpha)$ a heterogeneous parallel game with distribution function $\alpha$ such that $\alpha(0) = 0$ and $0$ is an accumulation point.
	If for any $i \in N$ it holds that $x_i(0) > 0$, then $x_i(t) > 0, \forall t \in \R_+^N$
\end{lemma}

\begin{proof}
	By the definition of an accumulation point, we get for $0$ in $\alpha$ that $\forall \delta > 0 \; \exists \epsilon > 0: p < \epsilon \Rightarrow \alpha(p) < \delta$.
	Therefore $\forall t \in \R_+^N$ if $\delta = \min_{j \in N \setminus \{i\}}\as^{(i, j)}(t)$ then $\exists \epsilon > 0: p < \epsilon \Rightarrow \alpha(p) < \as^{(i, j)}(t)$ for all $j \in N \setminus \{i\}$, which means that the lower latency link will always have flow some $\epsilon > 0$, and therefore $x_i(t) > 0$.
\end{proof}
Intuitively, the property describes the fact that having almost no toll fear makes a player not motivated to leave the edge, which in turn makes the network having little control over those players using toll regulations.
